---
title: "Project 2 Data Analytics"
author: "Skylar Furey"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, comment = NA, message = FALSE,fig.width=7, fig.height=4)
```

```{r}
Concussions <- read.csv("~/Desktop/Data Mining/STT450_R_Folder/Concussion.csv", header = TRUE)
```

## Simple Linear Regression Model #

```{r}
lm.fit =lm(Playtime_After~Playtime_Before, data= Concussions)
```

### Simple Linear Regression Model- Numerical Fit

```{r}
summary(lm.fit)
```

### Simple Linear Regression Model- Graphical Fit

```{r}
par(mfrow= c(2,2))
plot(lm.fit, col= 3)
```

```{r}
plot(Concussions$Playtime_Before, Concussions$Playtime_After, xlab= "Playtime Before", ylab= "Playtime After")
abline(lm.fit, lwd= 3, col= "red")
```

### Simple Linear Regression Model- ANOVA Table 

```{r}
anova(lm.fit)
```

## Interaction Terms #

```{r}
lm.fit2 =lm(Playtime_After~Playtime_Before * Games_Missed, data= Concussions)
```

### Interaction Terms- Numerical Fit 

```{r}
summary(lm.fit2)
```

### Interaction Terms- Graphical Fit 

```{r}
par(mfrow= c(2,2))
plot(lm.fit2, col=5)
```

### Interaction Terms- ANOVA Table

```{r}
anova(lm.fit2)
```

## Quadratic Linear Regression Model

```{r}
lm.fit3= lm(Concussions$Playtime_After~Concussions$Games_Missed)
lm.fit4= lm(Concussions$Playtime_After~poly(Concussions$Games_Missed, 2, raw=T))

```

### Quadratic Linear Regression Model- Numerical Fit

```{r}
summary(lm.fit4)
```

### Quadratic Linear Regression Model- Graphical Fit

```{r}
par(mfrow= c(2,4))
plot(lm.fit3, col=6)
plot(lm.fit4, col=7)
```

### Quadratic Linear Regression Model- ANOVA Table 

```{r}
anova(lm.fit3, lm.fit4)
```